---
title: "Day 1: MLM for Dyadic Data"
output:
  html_document: default
  pdf_document: default
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
```

[Back to schedule](schedule.html)

***

Read in the individual data (or a pairwise dataset) 

```{r}
library(tidyr)
library(dplyr)

acitelli_ind <- read.csv(file.choose(), header=TRUE)
```

Convert individual data to pairwise. If you imported a pairwise set, skip this chunk. 

```{r}
tempA <- acitelli_ind %>% 
  mutate(genderE = gender, partnum = 1) %>%
  mutate(gender = ifelse(gender == 1, "A", "P")) %>%
  gather(variable, value, self_pos:genderE) %>%
  unite(var_gender, variable, gender) %>%
  spread(var_gender, value)

tempB <- acitelli_ind %>% 
  mutate(genderE = gender, partnum = 2) %>%
  mutate(gender = ifelse(gender == 1, "P", "A")) %>%
  gather(variable, value, self_pos:genderE)%>%
  unite(var_gender, variable, gender) %>%
  spread(var_gender, value)

acitelli_pair <- bind_rows(tempA, tempB) %>%
  arrange(cuplid) 
  
rm(tempA, tempB)
```

#Multilevel Modeling (MLM) for Dyadic Data  

Now we're ready to do multilevel modeling with the pairwise dataset!

```{r}
#install.packages("nlme")
library(nlme)

mlm <- gls(satisfaction_A ~ genderE_A + Yearsmar,
          data = acitelli_pair,
          correlation = corCompSymm(form=~1|cuplid),
          na.action = na.omit)

summary(mlm)
```

##Interpretation
###Fixed Effects
**Intercept:**  Predicted level of satisfaction for people married about 11 years.  
**Effect of `genderE_A`:**  Husbands are more satisfied than wives by .027 units (not significant); we need to double because the difference between Husbands (+1) and Wives (-1) is two units.  
**Effect of `Yearsmar`:** For every year married, less satisfied by .0004 (not significant).

###Random Effects
`Rho` is the correlation of residuals, 0.62.  
`Residual standard error` is the error or unexplained variance (square-rooted).  

Partial ICC equals .620. Husbands and wives are very similar in their level of marital satisfaction.  
	
***  
	
[Back to schedule](schedule.html)

***  
