---
title: "Day 1: Actor-Partner Interdependence Model"
output:
  html_document: default
  html_notebook: default
---
Read in the individual data (or a pairwise dataset) 

```{r}
library(tidyr)
library(dplyr)
library(nlme)

acitelli_ind <- read.csv(file.choose(), header=TRUE)
```

Convert individual data to pairwise. If you imported a pairwise set, skip this chunk. I also create a gender variable that's a **factor** and has labels `hus` and `wife`. This vairable will be useful later. 

```{r}
tempA <- acitelli_ind %>% 
  mutate(genderE = gender, partnum = 1) %>%
  mutate(gender = ifelse(gender == 1, "A", "P")) %>%
  gather(variable, value, self_pos:genderE) %>%
  unite(var_gender, variable, gender) %>%
  spread(var_gender, value)

tempB <- acitelli_ind %>% 
  mutate(genderE = gender, partnum = 2) %>%
  mutate(gender = ifelse(gender == 1, "P", "A")) %>%
  gather(variable, value, self_pos:genderE)%>%
  unite(var_gender, variable, gender) %>%
  spread(var_gender, value)

acitelli_pair <- bind_rows(tempA, tempB) %>%
  arrange(cuplid) %>%
  mutate(gender_A = ifelse(genderE_A == 1, "hus", "wife"), gender_A = as.factor(gender_A)) 
  
rm(tempA, tempB)
```


#Indistinguishable dyads.

```{r}
apim_in <- gls(satisfaction_A ~ other_pos_A + other_pos_P,
          data = acitelli_pair,
          correlation = corCompSymm(form=~1|cuplid),
          na.action = na.omit)

summary(apim_in)
```

##Pseudo-R^2^
How much variance does teh actor and partner effects explain? First we run the empty model so that we can calculate the pseudo-R^2^.

```{r}
apim_in_empty <- gls(satisfaction_A ~ 1,
          data = acitelli_pair,
          correlation = corCompSymm(form=~1|cuplid),
          na.action = na.omit)

summary(apim_in_empty)
```

ICC = .618470
Residual SE^2 empty model = .246951
Residual SE^2 standard model = .174194
1 - (.174194 / .246951) = .295

Called the "pseudo R2"---29.5% of the variance is explained by other positivity of the actor and the partner. Set it to zero if it's negative.

##Interpretation
###Fixed Effects
Intercept:  Predicted level of satisfaction for those scoring zero on the actor and partner variables.  Because these variables are not centered, it is not all that meaningful.

OtherPos_A or the Actor Variable: If you see your partner positively, are you satisfied in the relation? Yes!

OtherPos_P or the Partner effect: If your partner sees you positively, are you satisfied in the relationship? (Or: If you see your partner positively, is your partner satisfied in the relationship?) Yes!

###Random Effects
"CSR diagonal" is the error or unexplained variance.
The partial ICC or CSR rho is .469.

The model explains nearly 30% of the variance. 


#Distingushable Dyads

##Interaction Approach
1. Add distinguishing variable as a covariate.  Note its coding.
2. Have the distinguishing variable interact with the actor and the partner effects.
3. These interactions evaluate whether actor and partner effects are the same for the two people.
4. Add a `weights =` argument to allow for different error variances for the two members.


```{r}
apim_di_int <- gls(satisfaction_A ~ other_pos_A + other_pos_P + genderE_A + other_pos_A*genderE_A + other_pos_P*genderE_A,
    data = acitelli_pair,
    correlation = corCompSymm(form=~1|cuplid), 
    weights = varIdent(form=~1|genderE_A), 
    na.action = na.omit)

summary(apim_di_int)
```

###Interpretation of Effects 

Intercept = .650854---The predicted score for husbands and wives who have a 0 on how positively they see the spouse (We should have centered!)
Gender_A  = .039605---Husband are very slightly more satisfied (about .08 points more) than wives when you control for how they both view their spouse.  (Recall wives are -1 on Gender_A and Husbands are +1; the difference between husbands and wives is then twice the difference of the effect of Gender_A.)
OtherPos_A = .401043---Actor Effect: The more positively you view your spouse, the more satisfied you are in the marriage.
OtherPos_P = .291564---Partner Effect:  The more positively your partner views you, the more satisfied you are in the marriage.
Gender_A*OtherPos_A = .023343---The actor effect is stronger for husbands.
Gender_A*OtherPos_P = -.029914---The partner effect is stronger H ??? W than W ??? H.

Actor Effect for Husbands = .401043 + .023343 = 0.424386
Actor Effect for Wives = .401043 - .023343  = 0.37770
Partner Effect for W ??? H = .291564 + (-.029914) = 0.261650
Partner Effect for H ??? W = .291564 - (-.029914) = 0.321478

.207460, error variance for Wives
.143139, error variance for Husbands



##Two-Intercept Approach
This involves a trick by which one equation becomes two.  We create two dummy variables: Hij which equals 1 for husbands and 0 for wives and Wij which equals 1 for wives and zero for husband.  We then estimate the following equation:

Yij = bHHij + aHHijAij + pHHijPij + Hijeij + bWWij + aWWijAij + pWWijPij + Wijeij

Note that the equation has no ordinary intercept, but rather, in some sense, two intercepts, bH and bW.  Note that when Hij = 1 and Wij = 0, the above becomes

Yij = bH + aHAij + pHPij + eij

and when Hij = 0 and Wij = 1, the above becomes

Yij = bW + aWAij + pWPij + eij

Thus, one equals becomes two and we have actor and partner for both members.

To implement this in R, we do the following.  
1. Add distinguishing variable as a factor, using `gender_A` created above.
2. Have no intercept in the fixed model by adding `-1` to the formula.
3. Have the distinguishing variable ('gender_A') interact with actor and partner effect, but no actor and partner main effects. We need to use `:` for this instead of `*`.
  i) Separate actor and partner effects will be estimated for each member.
4. Keep the `weights =` argument to allow for different error variances for the two members.


```{r}
apim_di_two <- gls(satisfaction_A ~ gender_A + other_pos_A:gender_A + other_pos_P:gender_A - 1,
    data = acitelli_pair,
    correlation = corCompSymm(form=~1|cuplid), 
    weights = varIdent(form=~1|genderE_A), 
    na.action = na.omit)

summary(apim_di_two)
```

We could also get pseudo R^2^ for husbands and wives separately. 

```{r}
apim_di_empty <- gls(satisfaction_A ~ gender_A - 1,
    data = acitelli_pair,
    correlation = corCompSymm(form=~1|cuplid), 
    weights = varIdent(form=~1|genderE_A), 
    na.action = na.omit)

summary(apim_di_empty)
```

Error variance for Wives: .280928 
- Pseudo R2 is 1 - .207460 /.280928 = .2615
Error variance for Husbands: .213248 
- Pseudo R2 is 1 - .143139/.213248 = .3288
