---
title: 'Day 1: Generalized Linear Mixed Models'
output:
  pdf_document: default
  html_document: default
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
```

[Back to schedule](schedule.html)

***  

Read in the individual data (or a pairwise dataset) 

```{r}
library(tidyr)
library(dplyr)

#install.packages("lme4")
library(lme4)

acitelli_ind <- read.csv(file.choose(), header=TRUE)
```

Convert individual data to pairwise. I also create a simhobs variable that will be our binary response, two dummy variables that will be useful for estimating separate random intercepts for men and women, and a count variable `cigarettes`. 

```{r}
tempA <- acitelli_ind %>% 
  mutate(genderE = gender, partnum = 1) %>%
  mutate(gender = ifelse(gender == 1, "A", "P")) %>%
  gather(variable, value, self_pos:genderE) %>%
  unite(var_gender, variable, gender) %>%
  spread(var_gender, value)

tempB <- acitelli_ind %>% 
  mutate(genderE = gender, partnum = 2) %>%
  mutate(gender = ifelse(gender == 1, "P", "A")) %>%
  gather(variable, value, self_pos:genderE)%>%
  unite(var_gender, variable, gender) %>%
  spread(var_gender, value)

acitelli_pair <- bind_rows(tempA, tempB) %>%
  arrange(cuplid) %>%
  mutate(gender_A = ifelse(genderE_A == 1, "hus", "wife"), 
         gender_A = as.factor(gender_A),
         simhob_bin_A = ifelse(simhob_A == 1, 1, 0), #forced binary variable
         man = ifelse(genderE_A == 1, 1, 0),
         woman = ifelse(genderE_A == 1, 0, 1),
         cigarettes_A = rpois(296, 0.7)) %>%
  group_by(cuplid) %>%
  mutate(cupcig = rpois(1, 0.7)) %>%
  ungroup(cuplid) %>%
  mutate(cigarettes_A = cigarettes_A + cupcig)
  
rm(tempA, tempB)
```

#Logistic Regression and Poisson Regression

First, let's only select the women so we can do away with the nesting. 

```{r}
acitelli_women <- acitelli_pair %>%
  filter(genderE_A == -1)
```

##Logistic Regression

Then we run a ordinary logistic regression. 

```{r}
logistic_reg <- glm(simhob_bin_A ~ other_pos_A + other_pos_P,
                    data = acitelli_women,
                    family = binomial)

summary(logistic_reg)
```

This gives us the increase in log odds of a woman having the same hobbies as her husband for every 1 unit increase in other positivity of the actor and partner other positivity.  

If we want to easily switch estimates from "log odds" to "odds," we use the following:

```{r}
exp(logistic_reg$coefficients)
```

##Poisson Regression (aka, Log-Linear Regression)

First, we run the model assuming the variance equals the mean. 

```{r}
poisson_reg <- glm(cigarettes_A ~ other_pos_A + other_pos_P,
                   data = acitelli_women,
                   family = poisson)

summary(poisson_reg)
exp(poisson_reg$coefficients)
```

But, we can also run a model relaxing the variance = mean assumption. That is, we can include an over-dispersion parameter with `family = quasipoisson`. 

```{r}
poissonOD_reg <- glm(cigarettes_A ~ other_pos_A + other_pos_P,
                     data = acitelli_women,
                     family = quasipoisson)

summary(poissonOD_reg)
exp(poissonOD_reg$coefficients)
```


#Logistic Multilevel Modeling (Binary variables)

To account for the nonindependence, we can make use of the `glmer()` function from the `lme4` package. Note that we are asking for the variance of intercepts across dyads, that is the random intercept in traditonal multilevel modeling. The `gls()` function in the `nlme` package does not have an option for specifying a link function (i.e., there is no `family =` option). The syntax of `glmer()` differs a bit from `gls()` in that the random effects are specified within the formula: `+ (1/cuplid)`.

##Indistinguishable Dyads

```{r}
apim_bin <- glmer(simhob_bin_A ~ other_pos_A + other_pos_P 
                  + (1|cuplid),
                  data = acitelli_pair,
                  family = binomial,
                  na.action = na.omit)

summary(apim_bin)
```

##Distinguishable Dyads

Interaction approach. 

```{r}
apim_bin_di <- glmer(simhob_bin_A ~ other_pos_A + other_pos_P + genderE_A 
                     + other_pos_A*genderE_A + other_pos_P*genderE_A 
                     + (man + woman - 1|cuplid),
                     data = acitelli_pair,
                     family = binomial,
                     na.action = na.omit)

summary(apim_bin_di)
```

Two-intercept model.

```{r}
apim_bin_di_two <- glmer(simhob_bin_A ~ gender_A + other_pos_A:gender_A + other_pos_P:gender_A - 1
                         + (man + woman - 1|cuplid),
                         data = acitelli_pair,
                         family = binomial,
                         na.action = na.omit)

summary(apim_bin_di_two)
```

#Log-Linear Multilevel Modeling (Count variables)

##Indistinguishable Dyads

```{r}
apim_poi <- glmer(cigarettes_A ~ other_pos_A + other_pos_P 
                  + (1|cuplid),
                  data = acitelli_pair,
                  family = poisson,
                  na.action = na.omit)

summary(apim_poi)
```

##Distinguishable Dyads

Interaction approach. 

```{r}
apim_poi_di <- glmer(cigarettes_A ~ other_pos_A + other_pos_P + genderE_A 
                     + other_pos_A*genderE_A + other_pos_P*genderE_A 
                     + (man + woman - 1|cuplid),
                     data = acitelli_pair,
                     family = poisson,
                     na.action = na.omit)

summary(apim_poi_di)
```

Two-intercept model.

```{r}
apim_poi_di_two <- glmer(cigarettes_A ~ gender_A + other_pos_A:gender_A + other_pos_P:gender_A - 1
                         + (man + woman - 1|cuplid),
                         data = acitelli_pair,
                         family = poisson,
                         na.action = na.omit)

summary(apim_poi_di_two)
```

#Generalized Estimating Equations (GEE)

##Indistinguishable Dyads

```{r}
#install.packages("gee")
library(gee)

apim_gee <- gee(simhob_bin_A ~ other_pos_A + other_pos_P, 
                id = cuplid,
                data = acitelli_pair, 
                na.action = na.omit,
                family = binomial, 
                corstr = "unstructured")

summary(apim_gee)
```

***

[Back to schedule](schedule.html)

*** 


